Competing Bandits:\\ The Perils of Exploration under Competition \thanks{This is an extended abstract of \citet{CompetingBandits-merged}, which in turn subsumes conference papers \citet{CompetingBandits-itcs18} and \citet{CompetingBandits-ec19}.}

### 2021-01-19 journal paper/content/sec-theory.tex
Space-saving
space-saving.

%\xhdr{Notation.} 
Let $\rew_i(n)$ denote the agent's realized reward observed by principal $i$ at local step $n$, \ie the reward collected by algorithm $\alg[i]$ in this local step. For a global round $t$, let $n_i(t)$ denote the number of global rounds before $t$ in which principal $i$ is chosen. Note that
    \[ \PMR_i(t):=  \E\sbr{ r_t\mid i_t = i }
        = \E\sbr{ \rew_i(n_i(t)+1) }. \]

@@

We characterize the inherent quality of an MAB algorithm in terms of its \emph{Bayesian Instantaneous Regret} (henceforth, \BIR), a standard notion from machine learning:
\begin{align}\label{eq:theory-BIR-defn}
\BIR_i(n) := \textstyle \E[\; \max_{a\in A} \mu_a - \rew_i(n) \;].
\end{align}
We are primarily interested in how fast \BIR decreases with $n$. (We treat the number of arms as a constant.) Intuitively, (much) better MAB algorithms tend to have a (much) smaller BIR, see Appendix~\ref{app:bg} for background.

An algorithm is called \emph{\bmonotone} if it can only get better over time, in the Bayesian sense: namely, if $\E[\rew_i(\cdot)]$ is non-decreasing, and therefore $\BIR(\cdot)$ is non-decreasing. This is a mild assumption, see
Appendix~\ref{app:examples}.

@@

\begin{theorem}\label{thm:random-clean}
Fix a \HardMaxRandom response function $\respF$. Suppose algorithms \alg[1], \alg[2] are \bmonotone and $T$-oblivious, and \eqref{eq:random-assn} holds. If \alg[1] \emph{BIR-dominates} \alg[2] starting from step $n_0$, then each agent $t\geq \max(n_0,m_0)$ chooses principal $1$ with probability $\respF(1) = 1- \eps_0$ (which is the largest possible selection probability for this response function).
\end{theorem}

@@

\begin{corollary}\label{cor:random}
Fix a \HardMaxRandom response function $\respF$. Consider the competition game in which all feasible MAB algorithms are $T$-oblivious, \bmonotone, and satisfy \eqref{eq:random-assn} for some fixed $m_0$. Suppose some feasible algorithm \alg  BIR-dominates all other feasible algorithms, starting from some local step $n_0$. Then, for any sufficiently large time horizon $T$, this game has a unique Nash equilibrium, in which both principals choose algorithm \alg.
\end{corollary}

>>>>

%\xhdr{Notation.}
Let $\rew_i(n)$ denote the agent's realized reward observed by principal $i$ at local step $n$, \ie the reward collected by algorithm $\alg[i]$ in this local step. For a global round $t$, let $n_i(t)$ denote the number of global rounds before $t$ in which principal $i$ is chosen. We will use the fact that
    $ \PMR_i(t):=  \E\sbr{ r_t\mid i_t = i }
        = \E\sbr{ \rew_i(n_i(t)+1) }. $

@@

We characterize the inherent quality of an MAB algorithm in terms of its \emph{Bayesian Instantaneous Regret} (henceforth, \BIR), a standard notion from machine learning:
\begin{align}\label{eq:theory-BIR-defn}
\BIR_i(n) := \textstyle \E[\; \max_{a\in A} \mu_a - \rew_i(n) \;].
\end{align}
We are primarily interested in how fast \BIR decreases with $n$. (We treat the number of arms as a constant.) Intuitively, (much) better MAB algorithms tend to have a (much) smaller BIR, see Appendix~\ref{app:bg} for background.
An algorithm is called \emph{\bmonotone} if it can only get better over time, in the Bayesian sense: namely, if $\E[\rew_i(\cdot)]$ is non-decreasing, and therefore $\BIR(\cdot)$ is non-decreasing. This is a mild assumption, see
Appendix~\ref{app:examples}.

@@

\begin{theorem}\label{thm:random-clean}
Fix a \HardMaxRandom response function $\respF$. Suppose algorithms \alg[1], \alg[2] are \bmonotone and $T$-oblivious, and \eqref{eq:random-assn} holds. If \alg[1] \emph{BIR-dominates} \alg[2] starting from step $n_0$, then each agent $t\geq \max(n_0,m_0)$ chooses principal $1$ with probability $\respF(1) = 1- \eps_0$ (which is the largest possible probability for this response function).
\end{theorem}

@@

\begin{corollary}\label{cor:random}
Fix a \HardMaxRandom response function $\respF$. Consider the competition game in which all feasible MAB algorithms are $T$-oblivious, \bmonotone, and satisfy \eqref{eq:random-assn} for some fixed $m_0$. Suppose some feasible algorithm \alg  BIR-dominates all other feasible algorithms, starting from some local step $n_0$. Then, for any sufficiently large time horizon $T$, this game has a unique Nash equilibrium: both principals choose algorithm \alg.
\end{corollary}
